---
title: "SafePaths SEIR Impact"
author: "Andrea Nuzzo"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: yes
  html_document: default
---

```{r message=FALSE, include=FALSE}
library(deSolve)
library(tidyverse)
library(ggthemes)
library(RCurl)
library(lubridate)
```

# Intro
Scope of this model is to demonstrate the impact of contact tracing procedures to limit the spreading of infectious diseases during outbreaks.

# Current situation as of `r Sys.Date()`

```{r}
# US population based on census = N
N = 329410596

JHU_Confirmed<- getURL('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv') %>% 
  read_csv() %>% 
  filter(`Country/Region`=='US') %>%
  gather(date, Infected, -`Province/State`, -`Country/Region`, -Lat, -Long) %>% 
  group_by(`Province/State`) %>% 
  mutate(date = mdy(date), 
         time_from_0 = date - min(date)) 

JHU_Recovered <- getURL('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv') %>% 
  read_csv() %>% 
  filter(`Country/Region`=='US') %>%
  gather(date, Recovered, -`Province/State`, -`Country/Region`, -Lat, -Long)%>% 
  group_by(`Province/State`) %>% 
  mutate(date = mdy(date), 
         time_from_0 = date - min(date))

JHU_Deaths <- getURL('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv') %>% 
  read_csv() %>% 
  filter(`Country/Region`=='US') %>%
  gather(date, Deaths, -`Province/State`, -`Country/Region`, -Lat, -Long) %>% 
  group_by(`Province/State`) %>% 
  mutate(date = mdy(date), 
         time_from_0 = date - min(date))

Current_scenario <- JHU_Confirmed %>% 
  group_by(date) %>% 
  summarize(Infected = sum(Infected)) %>% 
  full_join(JHU_Recovered %>% 
  group_by(date) %>% 
  summarize(Recovered = sum(Recovered)))%>% 
  full_join(JHU_Deaths %>% 
  group_by(date) %>% 
  summarize(Deaths = sum(Deaths))) %>% 
  ungroup%>% 
  mutate(time_from_0 = as.numeric(date - min(date)))

Current_scenario %>% 
  gather(Compartments, Cases, -date, -time_from_0) %>% 
    mutate(perc = Cases/N) %>% 
    ggplot(aes(x=time_from_0, y=perc, color=Compartments)) +
    # ylim(c(0,1.1))+
    geom_line()+
    theme_clean()+
    scale_color_manual(values = c(Infected = 'red', 
                                  Recovered = 'green', 
                                  Deaths = 'black')) 

Infected_symptomatic = tail(Current_scenario, 1)$Infected
Recovered = tail(Current_scenario, 1)$Recovered
Deaths = tail(Current_scenario, 1)$Deaths



```

# SEIR
This simulation is based on a SEIR model with the following modifications:

- Exposed *(include infected asymptomatic and non-infected contacts)* are subdivided into two mutually exclusive proportions: 
    - Followers: contacts who receive the notification of exposure risk _and_ self-isolate themselves from further contact
    - Non-followers: contacts who receive the notification of exposure risk _and_ self-isolate themselves from further contact

In the ideal scenario, the adoption of contact tracing practices will a) lower the contact rate per day and therefore b) lower the Exposed population

$T_{inf}$ = infectious period

$T_{lat}$ = latent_period

$c$ = contacts/day 

$p$ = probability of infection if exposed (estimated based on R0)

$\tau$ = percentage of non-followers among susceptible people

$\displaystyle \beta = c * p * \tau$

$\displaystyle \delta = \frac{1}{T_{inc}}$

$\displaystyle \gamma = \frac{1}{T_{lat}}$

$\displaystyle \mu$ = deaths/day (estimated)

$\displaystyle R_0 = \beta \gamma$ 

Then the model subdivides the last compartment into $Mild$, $Recovered$, $Dead$, with imputed percentages. 

## Data and assumptions 

```{r echo=TRUE}

# Estimated parameters
transmission_probability = 0.07
contact_rate = 22 
infectious_period = 2.9
latent_period = 5.2
non_symptomatic_ratio = 0.6
mortality_rate = 0.033
case_to_fatality = 0.02


# Current state as JHU dashboard for March 22
Exposed = (Infected_symptomatic/non_symptomatic_ratio)*contact_rate/as.integer(Sys.Date() - mdy('1/15/2020')) #Inferred based on assumptions
Susceptible = N - (Exposed + Infected_symptomatic/non_symptomatic_ratio + Infected_symptomatic + Recovered + Deaths)
obs_time = 300


## Derived values
Rt = contact_rate* infectious_period * transmission_probability
timepoints = seq(0, obs_time, by = .5)
```

## Model

$\displaystyle \frac{\partial{S}}{\partial{t}} = -\beta SI - \mu S$

$\displaystyle \frac{\partial{E}}{\partial{t}} =  \beta SI  - \delta E$

$\displaystyle \frac{\partial{I}}{\partial{t}} = \delta E - \gamma I_s$

$\displaystyle \frac{\partial{R}}{\partial{t}} = \gamma I_s$

```{r}

## Test values:
follower_percentage = 0.5

state_values <- c(
  S = Susceptible/N,
  E = Exposed/N,
  Is = Infected_symptomatic/N, 
  R = Recovered/N
)


SEIR.model <- function(current_timepoint, state_values, parameters){
  
  ## create state valriables
  S  = state_values[1] # susceptibles
  E  = state_values[2] # exposed
  Is = state_values[3] # infected symptomatic
  R  = state_values[4] # recovered
  
  with(
    as.list(parameters), 
    {
      ## compute derivatives
      dS = (-beta_value * S * Is)
      dE = ((beta_value * S * Is) - (delta_value * E))
      dI = (delta_value * E) - (gamma_value * Is)
      dR = (gamma_value * Is)
      
      # combine 
      results = c(dS, dE, dI, dR)
      list(results)
    }
  )
  
}

```


```{r}
for (followers_perc in seq(0.0, 0.99, by=.1)) {
  
  tau = (1-followers_perc)
  
  parameters = c(
  beta_value = contact_rate *tau * transmission_probability, 
  gamma_value = 1 / infectious_period, 
  delta_value = 1 / latent_period
)

  
  R0_est <- with(as.list(parameters), beta_value/gamma_value)
  
  out <- lsoda(y = state_values, times = timepoints, func = SEIR.model, parms = parameters)

  out <- out %>% 
    as.data.frame()
  
  p <- out %>% 
    gather(Compartments, perc, -time) %>% 
    ggplot(aes(x=time, y=perc, color=Compartments)) +
    ylim(c(0,1.1))+
    geom_line()+
    theme_clean()+
    scale_color_manual(values = c(S='grey', 
                                  E = 'blue',
                                  Ia = 'pink', 
                                  Is = 'red', 
                                  R = 'green', 
                                  D = 'black')) +
    labs(title=paste0("SEIsR model, ", followers_perc*100, "% adoption rate"),
         subtitle=paste0('Contact Rate = ', round(tau*contact_rate), '; Rt = ', round(R0_est, 2)))
    
  print(p)
  
}

```


